Effects of sexual selection on non-sexual fitness components: a meta-analysis

### 2018-03-06 meta_analysis_sexual_selection_pop_fitness.Rmd
6/3 Update with more supp tables

This latter criterion is likely to be contentious, because there is rarely enough data justify the assumption that a particular trait is (or is not) correlated with population fitness. We therefore relied on our best judgement when deciding which studies to exclude (see Table **XXX**). The inclusion/exlusion critera as applied to each study are detailed in **Table S1**.

**Table S1:** An eligibility criteria was based on four features a study needed to include (discussed above), to be eligable for inclusion in the meta-analysis the study needed to satisfy all criteria. Here we applied a step-wise process to the studies that had their full-text read and excluded them when they first failed to meet the criteria. Additional notes documenting reasons behind exclusion were also taken.
 <br/><br/> 
<input type=button class=hideshow></input>
```{r}
Eligibility.criteria <- read.csv('data/Eligibility Workbook(22.02).csv', 
                                 fileEncoding="UTF-8")
kable(Eligibility.criteria, "html") %>%
  kable_styling() %>%
  scroll_box(width = "800px", height = "500px")
```

@@

4. We also collected various covariates for some of the studies (**Table S2**), which are discussed later.

@@

**Table S2:** Table of effect sizes included in our meta-analysis. See the text following the table for an explanation of each column.
 <br/><br/> 
<input type=button class=hideshow></input>
```{r}
# Load the data and clean up the variable formats
prelim.data <- read.csv('data/Preliminary data frame 22.2.18.csv')
prelim.data$Study.ID <- prelim.data$Study.ID %>% factor()
prelim.data$Taxon <- prelim.data$Taxon %>% factor()
prelim.data$Group.ID <- prelim.data$Group.ID %>% factor()
prelim.data$Authors <- prelim.data$Authors %>% factor()
prelim.data$Environment <- prelim.data$Environment %>% factor %>% relevel(ref="Unstressed")
prelim.data$Sex <- prelim.data$Sex %>% factor %>% relevel(ref="B")
prelim.data$Ambiguous <- prelim.data$Ambiguous %>% factor()
prelim.data$Species <- prelim.data$Species %>% factor()
#Outcome.Class.2 is using the categories that were decided by survey. I am keeping both just to check them against each other (how much of a difference it makes)
prelim.data$Outcome.Class <- prelim.data$Outcome.Class %>% factor() %>% relevel(ref="Indirect")
prelim.data$Enforced.Monogamy <- prelim.data$Enforced.Monogamy %>% factor() %>% relevel(ref="NO")
prelim.data$Pre.cop <- prelim.data$Pre.cop %>% factor() %>% relevel(ref="0")
prelim.data$Post.cop <- prelim.data$Post.cop %>% factor() %>% relevel(ref="0")
prelim.data$Blinding <- prelim.data$Blinding %>% factor()

@@

**Study ID**: An ID given to the published paper the effect size is sourced from

**Group ID**: An ID given to the research group that may have published several papers on the same species usuing the same or very similar experimental setup

**Species and Species ID**: Same thing

@@

**Post cop and Pre cop**: Whether a study allowed Pre/Post-copulatory sexual selection (1) or not (0)

**Blind**: Whether the study was blind or not

**Generations** The number of generations run before fitness outcomes were measured

@@

**Sex**: Whether the fitness outcome was measured for females (F), males (M) or both (B). Studies that reported 'both' would pool results from males and females.

@@

**Outcome Class**: Grouped as Direct, Indirect or Ambiguous.

@@

@@

**JIF**: Journal Impact factor at year of publication. Several impact factors were unable to be determined/found and are NA.
<br></br>

@@

**Table S3:** Table of effect sizes included in our meta-analysis. See the text following the data table for an explanation of each column.
 <br/><br/> 
<input type=button class=hideshow></input>
```{r message=FALSE, warning=FALSE}
n.blind.ones <- (sum(prelim.data$Blind == "Blind"))
prelim.data %>% 
  summarise(
    Effect_sizes_.Totalq = n(), 
    Publications = prelim.data$Study.ID %>% unique() %>% length(),
    Blind_experiments = n.blind.ones,
    Effect_sizes_.Stressedq = (sum(prelim.data$Environment == "Stressed")),
    Effect_sizes_.Unstressedq = (sum(prelim.data$Environment == "Unstressed")),         
    Effect_sizes_.Maleq = (sum(prelim.data$Sex == "M")),
    Effect_sizes_.Femaleq = (sum(prelim.data$Sex == "F")),
    Effect_sizes_.Both_sexesq = (sum(prelim.data$Sex == "B")),
    Different_species =  prelim.data$Species %>% unique() %>% length(),
    Effect_sizes_.Beetleq = sum(Taxon == "Beetle"),
    Effect_sizes_.Flyq = sum(Taxon == "Fly"),
    Effect_sizes_.Mouseq = sum(Taxon == "Mouse"),
    Effect_sizes_.Nematodeq = sum(Taxon == "Nematode"),
    Effect_sizes_.Miteq = sum(Taxon == "Mite"),
    Effect_sizes_.Cricketq = sum(Taxon == "Cricket"),
    Effect_sizes_.Guppyq = sum(Taxon == "Guppy")) %>% melt() %>%
  mutate(variable = gsub("_", " ", variable),
         variable = gsub("[.]", "(", variable),
         variable = gsub("q", ")", variable)) %>% 
  rename_("n" = "value", " " = "variable") %>% 
  pander(split.cell = 40, split.table = Inf)
```
 <br/><br/> 
**Table S4:** Table of fitness outcomes included in our meta-analysis by sex.
 <br/><br/> 
<input type=button class=hideshow></input>
```{r}
Outcome_and_sex <- as.data.frame.matrix(table(prelim.data$Outcome, prelim.data$Sex))
colnames(Outcome_and_sex) <- cbind("Both", "Male", "Female")

@@

@@

```{r}
#run model without taxon: same random effects as above
model.complete2 <- rma.mv(g, var.g, 
                          mods = ~ 1 + Sex + Environment + Sex:Environment, 
                          random = list(~ 1 | Study.ID, 
                                       ~ 1 | Outcome), 
                          method = "REML", 
                          data = restricted.data)
summary(model.complete2)
```

@@

>>>>

This latter criterion is likely to be contentious, because there is rarely enough data justify the assumption that a particular trait is (or is not) correlated with population fitness. We therefore relied on our best judgement when deciding which studies to exclude (see **Table S1**). The inclusion/exlusion critera as applied to each study are detailed in **Table S2**.

**Table S1:** We classed each of the twenty fitness related outcomes into three broad groups of direct, indirect and ambiguous based on the established link with population fitness, the directionality of the measure. Here we detailed how these outcomes were measured in the studies of this meta-analysis. In the accompanying box we provide a legend to the references cited in the table.

 <br/><br/>
```{r, warning=FALSE}
outcome.descriptions <- read.csv('data/outcome.descriptions.csv', 
                                 fileEncoding="UTF-8")
kable(outcome.descriptions, "html") %>%
  kable_styling() %>%
  scroll_box(width = "800px", height = "500px")
```
  <br/><br/> 
<input type=button class=hideshow></input>
```{r, warning=FALSE}
outcome.references <- read.csv('data/references.tableS1.csv', 
                                 fileEncoding="UTF-8")
kable(outcome.references, "html") %>%
  kable_styling() %>%
  scroll_box(width = "800px", height = "250px")
```
  <br/><br/>

**Table S2:** An eligibility criteria was based on four features a study needed to include (discussed above), to be eligable for inclusion in the meta-analysis the study needed to satisfy all criteria. Here we applied a step-wise process to the studies that had their full-text read and excluded them when they first failed to meet the criteria. Additional notes documenting reasons behind exclusion were also taken.
 <br/><br/> 
<input type=button class=hideshow></input>
```{r}
Eligibility.criteria <- read.csv('data/Eligibility Workbook(22.02).csv', 
                                 fileEncoding="UTF-8")
kable(Eligibility.criteria, "html") %>%
  kable_styling() %>%
  scroll_box(width = "800px", height = "500px")
```

@@

4. We also collected various covariates for some of the studies (**Table S3**), which are discussed later.

@@

**Table S3:** Table of effect sizes included in our meta-analysis. See the text following the table for an explanation of each column.
 <br/><br/> 
<input type=button class=hideshow></input>
```{r}
# Load the data and clean up the variable formats
prelim.data <- read.csv('data/Preliminary data frame 22.2.18.csv')
prelim.data$Study.ID <- prelim.data$Study.ID %>% factor()
prelim.data$Taxon <- prelim.data$Taxon %>% factor()
prelim.data$Group.ID <- prelim.data$Group.ID %>% factor()
prelim.data$Authors <- prelim.data$Authors %>% factor()
prelim.data$Environment <- prelim.data$Environment %>% factor %>% relevel(ref="Unstressed")
prelim.data$Sex <- prelim.data$Sex %>% factor %>% relevel(ref="B")
prelim.data$Ambiguous <- prelim.data$Ambiguous %>% factor()
prelim.data$Species <- prelim.data$Species %>% factor()
#Outcome.Class.2 is using the categories that were decided by survey. I am keeping both just to check them against each other (how much of a difference it makes)
prelim.data$Outcome.Class <- prelim.data$Outcome.Class %>% factor() %>% relevel(ref="Indirect")
prelim.data$Enforced.Monogamy <- prelim.data$Enforced.Monogamy %>% factor() %>% relevel(ref="NO")
prelim.data$Pre.cop <- prelim.data$Pre.cop %>% factor() %>% relevel(ref="0")
prelim.data$Post.cop <- prelim.data$Post.cop %>% factor() %>% relevel(ref="0")
prelim.data$Blinding <- prelim.data$Blinding %>% factor()

@@

**Study ID**: An ID given to the published paper the effect size is sourced from (n=65).

**Group ID**: An ID given to the research group that may have published several papers on the same species usuing the same or very similar experimental setup. [Was not use in analysis]

**Species**: The species used in the experimental evolution procedure (n = 15).

**Taxon**: The taxon to which the species belongs. One of the following: Beetle, fly, mouse, nematode, guppy, mite and cricket (taxa were selected arbitrarily based on the available data).

@@

**Post cop and Pre cop**: Whether a study allowed Pre/Post-copulatory sexual selection (1) or not (0).

**Blinding**: A binary classification, describing whether blind protocols were used during the experiment. Papers were assumed to be not blind unless declared otherwise.

**Generations**: The number of generations that the species was subject to differing levels of sexual selection, ranging from 1 to 162.

@@

**n**: Pooled sample size of the paired treatments.

**Outcome**: The fitness related outcome that was measured, e.g. fecundity, survival, or mating success (see Table S1 for all 20 categories). We applied our own classifications rather than relying on those provided by the authors, because different papers sometimes used different names for the same trait.

**Outcome Class**: To help guide analysis the outcomes were classed into three categories; ambiguous, indirect and direct (see Table S1).

**Sex**: A moderator variable with three levels, describing whether the effect size in question comes from a measurement of males (M), females (F), or individuals of both sexes (B).

@@

@@

**Positive Fitness**: Whether the measurment used in the study is beneficial for fitness (1) or not (0). Note that g has already been multiplied by this column.

@@

**JIF**: Journal Impact factor at year of publication. Several impact factors were unable to be determined/found and are NA.We obtained the journal impact factor for each effect size at the time of publication using InCites Journal Citation Reports.
<br></br>

@@

**Table S4:** Table of effect sizes included in our meta-analysis. See the text following the data table for an explanation of each column.
 <br/><br/> 
<input type=button class=hideshow></input>
```{r message=FALSE, warning=FALSE}
n.blind.ones <- (sum(prelim.data$Blind == "Blind"))
prelim.data %>% 
  summarise(
    Effect_sizes_.Totalq = n(), 
    Publications = prelim.data$Study.ID %>% unique() %>% length(),
    Blind_experiments = n.blind.ones,
    Effect_sizes_.Stressedq = (sum(prelim.data$Environment == "Stressed")),
    Effect_sizes_.Unstressedq = (sum(prelim.data$Environment == "Unstressed")),         
    Effect_sizes_.Maleq = (sum(prelim.data$Sex == "M")),
    Effect_sizes_.Femaleq = (sum(prelim.data$Sex == "F")),
    Effect_sizes_.Both_sexesq = (sum(prelim.data$Sex == "B")),
    Different_species =  prelim.data$Species %>% unique() %>% length(),
    Effect_sizes_.Beetleq = sum(Taxon == "Beetle"),
    Effect_sizes_.Flyq = sum(Taxon == "Fly"),
    Effect_sizes_.Mouseq = sum(Taxon == "Mouse"),
    Effect_sizes_.Nematodeq = sum(Taxon == "Nematode"),
    Effect_sizes_.Miteq = sum(Taxon == "Mite"),
    Effect_sizes_.Cricketq = sum(Taxon == "Cricket"),
    Effect_sizes_.Guppyq = sum(Taxon == "Guppy")) %>% melt() %>%
  mutate(variable = gsub("_", " ", variable),
         variable = gsub("[.]", "(", variable),
         variable = gsub("q", ")", variable)) %>% 
  rename_("n" = "value", " " = "variable") %>% 
  pander(split.cell = 40, split.table = Inf)
```
 <br/><br/> 
**Table S5:** Table of fitness outcomes included in our meta-analysis by sex.
 <br/><br/> 
<input type=button class=hideshow></input>
```{r}
Outcome_and_sex <- as.data.frame.matrix(table(prelim.data$Outcome, prelim.data$Sex))
colnames(Outcome_and_sex) <- cbind("Both", "Female", "Male")

@@

###Quantitative Summary Mean

We can obtain a bootstrap weighted mean for the entire dataset. Although this does not take into account any moderator variables it does provide a rough guide on where most of the datapoints sit:

```{r}
#Run function
get.CIs.for.weighted.mean <- function(x, weights){
  total.resamples <- 10000
  num.chunks <- ceiling(length(x)*total.resamples / 10^7) # To be nice to RAM, let's keep these to ten million or less 
  if(num.chunks == 1){
    rand <- sample(length(x), length(x)*total.resamples, replace = TRUE)
    resampled.weighted.means <- data.frame(x = x[rand], w = weights[rand], replicate = rep(1:total.resamples, each = length(x))) %>% 
      group_by(replicate) %>% 
      summarise(w.mean = weighted.mean(x, w)) %>% .$w.mean
  }

  else{
    resampled.weighted.means <- vector(mode = "list", length = 0)
    number.of.resamples <- floor(total.resamples/num.chunks)
    final.chunk.size <- total.resamples - number.of.resamples * (num.chunks-1)
    for(i in 1:num.chunks){
      if(i == num.chunks) number.of.resamples <- final.chunk.size
      rand <- sample(length(x), length(x) * number.of.resamples, replace = TRUE)
      resampled.weighted.means[[i]] <- data.frame(x = x[rand], w = weights[rand], replicate = rep(1:number.of.resamples, each = length(x))) %>% 
        group_by(replicate) %>% 
        summarise(w.mean = weighted.mean(x, w)) %>% .$w.mean
    }
    resampled.weighted.means <- unlist(resampled.weighted.means)
  }

  data.frame(w.mean = median(resampled.weighted.means),
             lowerCI = quantile(resampled.weighted.means, probs = 0.025) %>% as.numeric(),
             upperCI = quantile(resampled.weighted.means, probs = 0.975) %>% as.numeric())
}

#Get weights
prelim.data$Standard.weights <- (1/prelim.data$var.g)

#Run function with weights
get.CIs.for.weighted.mean(prelim.data$g, prelim.data$Standard.weights) %>% pander()
```

>Could this be used as a line in the forest plot showing the average (or as a polygon at the bottom)

@@

```{r}
#run model without taxon: same random effects as above
model.complete2 <- rma.mv(g, var.g, 
                          mods = ~ 1 + Sex * Environment, 
                          random = list(~ 1 | Study.ID, 
                                       ~ 1 | Outcome), 
                          method = "REML", 
                          data = restricted.data)
summary(model.complete2)
```

@@

Using an anova we can test to see if a female in a stressed environment gains significantly greater fitness benefits than in an unstressed environment:

```{r}
anova(model.complete2, L=c(0, 0, 1, -1, 0, 0)) 
```

```{r}
anova(model.complete2, L=c(0, 1, 0, -1, 0, 0)) 
```
